A multivariate skew-normal finite mixture model for analysis of infant development trajectories

### 2019-05-23 manuscript/manuscript.tex
appendix

@@

\subsection{Simulation to Assess Sensitivity to Mispecified H}

@@

@@

@@

    \item $\mathbf{Y}$: A $n \times p$ matrix containing all multivariate skew-normal regression outcomes such that $y_{ij}$ is the $j^{th}$ outcome observed for subject $i$, where $i = 1,...,n$ and $j = 1,...p$.

    \item $\mathbf{X}$: A $n \times m$ matrix containing all multivariate skew-normal regression covariates such that $x_{ij}$ is the $j^{th}$ covariate value for subject $i$.

@@

    \item $\mathbf{t}$: A $n \times 1$ vector of truncated normal random effects used in the stochastic representation of the multivariate skew-normal distribution. For $i = 1,...,n$, $t_i \stackrel{iid}{\sim}T_{[0,\infty)(0,1)}$

@@

Without loss of generality, we derive the full conditional distributions for the multivariate skew-normal regression model component under the assumption that all observations belong to a single cluster. To make the extension to the case where more than one cluster is specified, simply apply these distributional forms to cluster specific parameters and data.

The multivariate skew-normal regression model can be written as follows. 
$$\mathbf{Y} = \mathbf{X}\mathbf{B} + \mathbf{t} \boldsymbol\psi^T + \mathbf{E} = \mathbf{X}^* \mathbf{B}^* + \mathbf{E}$$

For convenience, we define $\mathbf{X}^*$ as a $n \times (m + 1)$ matrix constructed by column binding $\mathbf{t}$ to $\mathbf{X}$, and $\mathbf{B}^*$ as a $(m+1) \times p$ matrix constructed by row binding $\boldsymbol\psi^T$ to $\mathbf{B}$. We assume that $t_i \stackrel{iid}{\sim}T_{[0,\infty})(0,1)$ and that $\mathbf{E}$ is made of row vectors $\boldsymbol\epsilon_i = (\epsilon_{i1},...,\epsilon_{ip})$ for $i = 1,...,n$, where $ \boldsymbol\epsilon_i \stackrel{iid}{\sim} N_p(0, \boldsymbol\Sigma)$.

@@

We choose conjugate priors for $\mathbf{B}^*$ and $\boldsymbol\Sigma$ as follows. 
$$\boldsymbol\Sigma \sim \text{inverse-Wishart}(\mathbf{V}_0,\nu_0) $$
$$\mathbf{B}^*|\boldsymbol\Sigma \sim MN_{(m+1)\times p}(\mathbf{B}_0^*,\mathbf{L}_0^{-1},\boldsymbol\Sigma)$$

We now dervive the joint posterior distribution of the parameters $\mathbf{B}^*$ and $\boldsymbol\Sigma$.
\begin{align*} 
p(\mathbf{B}^*,\boldsymbol\Sigma|\mathbf{X}^*,\mathbf{Y}) & \propto  p(\mathbf{Y}|\mathbf{X}^*,\mathbf{B}^*,\boldsymbol\Sigma)p(\mathbf{B}^*|\boldsymbol\Sigma)p(\boldsymbol\Sigma)\\
&  \propto |\boldsymbol\Sigma|^{-n/2}\exp \left \{ -\frac{1}{2} \tr((\mathbf{Y} - \mathbf{X^*}\mathbf{B}^*)^T(\mathbf{Y} - \mathbf{X^*}\mathbf{B}^*)\boldsymbol\Sigma^{-1})\right \} \\
& \times |\boldsymbol\Sigma|^{-m/2} \exp \left \{ -\frac{1}{2} \tr((\mathbf{B}^* - \mathbf{B}^*_0)^T \mathbf{L}_0(\mathbf{B}^* - \mathbf{B}^*_0)\boldsymbol\Sigma^{-1})\right \} \\
& \times |\boldsymbol\Sigma|^{(\nu_0 + p + 1)/2} \exp \left \{ -\frac{1}{2} \tr (\mathbf{V}_0 \boldsymbol\Sigma^{-1}) \right \}
\end{align*}

>>>>

\item Emphasize that PG data augmentation for the multinomial model results in a PG mixure of experts model, which is a computationally efficient way to model edge weights.

@@

\subsection{Simulation to Assess Sensitivity to Mispecified K}

@@

\item Include both time varying and non-time varying covariates for the within cluster covariate set.

@@

\item Discuss how we handle non-ignorable missingness

@@

    \item $\mathbf{Y}$: A $N \times J$ matrix containing all multivariate skew-normal  outcomes such that $y_{ij}$ is the $j^{th}$ outcome observed for subject $i$, where $i = 1,...,n$ and $j = 1,...J$.

    \item $\mathbf{X}$: A $n \times p$ matrix containing all multivariate skew-normal regression covariates such that $x_{ij}$ is the $j^{th}$ covariate value for subject $i$.

@@

    \item $\mathbf{t}$: An $n \times 1$ vector of truncated normal random effects used in the stochastic representation of the multivariate skew-normal distribution. For $i = 1,...,n$, $t_i \stackrel{iid}{\sim}T_{[0,\infty)}(0,1)$

@@

Without loss of generality, we derive the full conditional distributions for the multivariate skew-normal regression model component under the assumption that all observations belong to a single cluster. To make the extension to the case where more than one cluster is specified, simply apply these distributional forms to cluster specific parameters and data. Finally, we assume for the moment that we have complete data for all outcomes for each subject. We extend consider the case of missing data in section (INSERT SECTION).

The multivariate skew-normal regression model can be written as follows in matrix form. 
$$\mathbf{Y} = \mathbf{X}\mathbf{B} + \mathbf{t} \boldsymbol\psi^T + \mathbf{E} = \mathbf{X}^* \mathbf{B}^* + \mathbf{E}$$
The matrix $\mathbf{Y}$ is of dimension $n \times p$. For convenience, we define $\mathbf{X}^*$ as a $n \times (m + 1)$ matrix constructed by column binding $\mathbf{t}$ to $\mathbf{X}$, and $\mathbf{B}^*$ as a $(m+1) \times p$ matrix constructed by row binding $\boldsymbol\psi^T$ to $\mathbf{B}$. We assume that $t_i \stackrel{iid}{\sim}T_{[0,\infty})(0,1)$ and that $\mathbf{E}$ is made of row vectors $\boldsymbol\epsilon_i = (\epsilon_{i1},...,\epsilon_{ip})$ for $i = 1,...,n$, where $ \boldsymbol\epsilon_i \stackrel{iid}{\sim} N_p(0, \boldsymbol\Sigma)$.

@@

We choose conjugate priors for $\mathbf{B}^*$ and $\boldsymbol\Sigma$ as follows. 
$$\boldsymbol\Sigma \sim \text{inverse-Wishart}(\mathbf{V}_0,\nu_0)$$

$$\mathbf{B}^*|\boldsymbol\Sigma \sim MatNorm_{(m+1), p}(\mathbf{B}_0^*,\mathbf{L}_0^{-1},\boldsymbol\Sigma)$$

We now dervive the joint posterior distribution of the parameters $\mathbf{B}^*$ and $\boldsymbol\Sigma$.
\begin{align*} 
p(\mathbf{B}^*,\boldsymbol\Sigma|\mathbf{X}^*,\mathbf{Y}) & \propto  p(\mathbf{Y}|\mathbf{X}^*,\mathbf{B}^*,\boldsymbol\Sigma)p(\mathbf{B}^*|\boldsymbol\Sigma)p(\boldsymbol\Sigma)\\
&  \propto |\boldsymbol\Sigma|^{-n/2}\exp \left \{ -\frac{1}{2} \tr \left [(\mathbf{Y} - \mathbf{X^*}\mathbf{B}^*)^T(\mathbf{Y} - \mathbf{X^*}\mathbf{B}^*)\boldsymbol\Sigma^{-1} \right ]\right \} \\
& \times |\boldsymbol\Sigma|^{-(m+1)/2} \exp \left \{ -\frac{1}{2} \tr \left [(\mathbf{B}^* - \mathbf{B}^*_0)^T \mathbf{L}_0(\mathbf{B}^* - \mathbf{B}^*_0)\boldsymbol\Sigma^{-1} \right ]\right \} \\
& \times |\boldsymbol\Sigma|^{(\nu_0 + p + 1)/2} \exp \left \{ -\frac{1}{2} \tr (\mathbf{V}_0 \boldsymbol\Sigma^{-1}) \right \}
\end{align*}
